Exploring the Relationship Between β-carotene Levels, BCO1 Variants, and Atherosclerosis Progression through Data Analysis

### 2024-04-11 R/eda-code/exploratory-analysis.qmd
Part 4 edits

# Plot the data
library(ggplot2)
ggplot(body_weights_long, aes(x = Week, y = BodyWeight, color = Category)) +
  geom_line() +
  labs(title = "Comparison of Body Weight Changes Across Categories",
       x = "Week", y = "Body Weight") +
  theme_minimal()
```

>>>>

# Plot the data
library(ggplot2)
ggplot(body_weights_long, aes(x = Week, y = BodyWeight, color = Category)) +
  geom_line() +
  labs(title = "Comparison of Body Weight Changes Across Categories",
       x = "Week", y = "Body Weight") +
  theme_minimal()
figure_file2 = here("results", "figures", "category_distro.png")
ggsave(filename = figure_file2, plot=bar1) 
```

### 2024-04-11 R/processing-code/processingfile-v1.qmd
Part 4 edits

```{r}
dir.create("/Users/mutsa_n/Desktop/MADA-course/Nyamuranga-MADA-project/data/processed_data/", recursive = TRUE)
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(figdata, file = save_data_location)
```

>>>>

```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(figdata, file = save_data_location)
```

### 2024-04-11 products/manuscript/Manuscript.qmd
Part 4 edits

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

@@

# Load necessary package
library(readxl)

# Read Excel file
rawdata <- read_excel("~/Desktop/MADA-course/Nyamuranga-MADA-project/data/raw-data/elife-87430-data1-v1.xlsx")
```

@@

## Data aquisition

@@

## Data import and cleaning

@@

@@

@@

*Summarize what you did, what you found and what it means.*

>>>>

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(tidyr)
library(broom)
```

@@

# Load necessary package
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data/raw-data/elife-87430-data1-v1.xlsx")
rawdata <- readxl::read_excel(data_location)
```

@@

## Data Acquisition

@@

## Data Import and Cleaning

@@

### Time Series Regression Analysis
A time series regression analysis was conducted to investigate temporal trends and relationships between variables over time. The time series data were examined for stationarity and seasonality using appropriate statistical tests. The primary variables of interest were body weight over weeks. A linear regression model was fitted to the time series data to quantify the relationship between the predictor variables and the outcome variable over the study period. Model diagnostics, including residual analysis and goodness-of-fit measures, were performed to assess model adequacy.

### Basic Linear Regression Analysis
A basic linear regression analysis was performed to examine the association between predictor variables and the outcome variable. The predictor variables included control, baseline and beta-carotene, while the outcome variable was body weight. Prior to analysis, multicollinearity diagnostics were conducted to identify and mitigate issues arising from collinear predictor variables. The linear regression model was fitted to the data, and model assumptions were assessed through residual analysis and goodness-of-fit tests.

### Train/Test Linear Regression Analysis
To assess the generalizability and predictive performance of the linear regression model, a train/test approach was employed. The dataset was randomly divided into a training set (80% of the data) and a testing set (remaining 20 of the data). The linear regression model was fitted to the training set, and model performance was evaluated using the testing set. Performance metrics such as mean squared error (MSE) and R-squared were calculated to assess the predictive accuracy of the model.

### Statistical Software
All statistical analyses were conducted using [name of statistical software or programming language, e.g., R, Python]. Statistical significance was determined at the conventional alpha level of 0.05.

@@

```{r}
#| label: Full Model Fit
#| tbl-cap: "Full model fit table."
#| echo: FALSE
resulttable3 = readRDS(here("results","output","resulttable3.rds"))

print(resulttable3)
```
The predictive model demonstrated favorable performance, with an RMSE of 1.564 and an R-squared value of 0.954. The low RMSE indicates minimal deviation between predicted and actual body weight values, suggesting the model's predictive accuracy. Furthermore, the high R-squared value indicates that the model explains approximately 95.4% of the variance in body weight, indicating strong explanatory power. Interpretation of model coefficients revealed insights into the relative importance of predictors in influencing body weight.

The results highlight the effectiveness of the predictive model in capturing the relationship between predictors and body weight. The preprocessing steps, including dummy encoding, centering, and scaling, played a crucial role in enhancing model performance by standardizing and transforming predictors. The high explanatory power of the model suggests that the included predictors effectively explain variations in body weight.

@@
